Introducing ICBe: Very High Recall and Precision Event Extraction from Narratives about International Crises

### 2022-08-10 README.Rmd
readme update

  - [ICBe_V1_long_agreement.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds)
      + All coded values and information about how often they were selected by coders.  
  - [ICBe_V1_long_agreed.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_long_agreed.Rds)
      + `ICBe_V1_long_agreement.Rds` filtered down only to those codings that were agreed upon (see Algorithm 1 in paper).
  - [ICBe_V1_wide_agreed.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds)
      + `ICBe_V1_long_agreed.Rds` in wide form where each row is an event.

@@

### Self Contained Package

All of the files used to create the datasets, tables, figures.

## RMarkdown files

### Replication Data

 - 01_compile_saves_and_align
    + compiles the original coding files into `./replication_data/in/icb_coder_crisis_sentence_event.Rds`. The original coding files are not on the public repository. Public users will load `icb_coder_crisis_sentence_event.Rds` directly.
    + aligns codings from multiple GUI versions on similar source sentences
    + creates `./replication_data/temp/ICBe_V1_long_unclean.Rds.Rds` (not included in online repo)
 - 02_format_and_clean
    + applies cleaning dictionaries to create `./replication_data/temp/ICBe_V1_long_clean.Rds` and `./replication_data/temp/ICBe_V1_long_clean.Rds` (not included in online repo)
 - 03_aggregation
    + applies aggregation algorithm to create `./replication_data/out/ICBe_V1_long_agreement.Rds`, `./replication_data/out/ICBe_V1_long_agreed.Rds`, `./replication_data/out/ICBe_V1_wide_agreed.Rds`.

### Replication Paper

  - 04_ICBEdataset_paper_litreviewtable
      + creates the table `/tables/litreview_ft.Rds`.
  - 05_ICBEdataset_paper_PlotInterCoderAgreement
      + creates the figure `figures/p_percent_chose_tag_by_concept.png`
  - 06_ICBEdataset_figure_metro_maps
      + creates the metroplot Rds files in `/figures/metro_plots/.`.
  - 07_ICBEdataset_figure_semantic_embeddings
      + creates the figure `/figures/p_semantic_embeddings.png` and the `/tables/codings_wide_agreed_embeded_stratified_sample_ft.Rds` and `/tables/codings_wide_agreed_embeded_stratified_sample_ft_smaller.Rds`
  - 08_CaseStudy196
      + creates Cuban Missile Crisis comparison table `/tables/CaseStudy196_ft.Rds`. 
  - 09_CaseStudy426
      + creates DRC Civil War comparison table `/tables/CaseStudy426_ft.Rds`. 
  - 10_ICBEdataset_paper_figure_prep
      + creates the metroplot pngs in `/figures/.`.
  - 11_ICBEdataset_paper_appendixICBCrises
      +  creates `/tables/ft_crisis_text_counts.Rds`.
  - 12_ICBEdataset_paper_appendixVerbMeanings
      + creates `/tables/verbs_sentence_wordnet_top_ft.Rds`

@@

### External datasets
  - [The ICB project](https://sites.duke.edu/icbdata/) 
    + System-level (./data/icb1v14.Rds) and Actor-level (./data/icb2v14.Rds) datasets
    + Dyadic-Level Crisis Data ([source](https://sites.duke.edu/icbdata/data-collections/))
  - [Militarized Interstate Disputes (MID)](https://correlatesofwar.org/data-sets/MIDs) version 5.01 at the incident level (MIDI 5.01) and incident-participant level (MIDIP 5.01) converted to Rds. 
  - [UCDP Georeferenced Event Dataset (GED) Global version 21.1](https://ucdp.uu.se/downloads/index.html#ged_global)
  - cameo.eventcode.txt adapted from [CAMEO Conflict and Mediation Event Observations Codebook](https://parusanalytics.com/eventdata/cameo.dir/CAMEO.09b6.pdf)
  - [Phoenix Event data](https://databank.illinois.edu/datasets/IDB-2796521)
  - [Terrier event data](https://osf.io/4m2u7/files/) 
    + too large to include in the github repository
    + to replicate, download the folder 'largegeolocatedata' to ICBEdata/replication_paper/data/ignore and decompress
  - [ICEWS data](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/28075&version=30.0)
    + too large to include in the github repository
    + to replicate, download folder 'dataverse_files' to ICBEdata/replication_paper/data/ignore/ and decompress

>>>>

  - [ICBe_V1_long.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_long_agreed.Rds)
  - [ICBe_V1.1_long_agreement.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds)
      + All coded values individually along with information about how often they were selected by coders.  
      + `ICBe_V1_long.Rds` filtered down only to those codings that were agreed upon (see Algorithm 1 in paper).
  - [ICBe_V1.1_events_agreed_long.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds)
    - [ICBe_V1.1_events_agreed.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed_long.Rds)

      + `ICBe_V1.1_events_agreed.Rds` in wide form where each row is an event.

@@

A description of the file and folders in the repository used to create the datasets, tables, figures.

## data preparation

### replication_corpus

 - download_and_clean
    + creates a succinct rds of the crisis narratives: `.replication_corpus/data/out/icb_corpus_V1.0_May_16_2022`

### replication_data

 - 01_compile_saves_and_align
    + compiles the original coding files into `./replication_data/in/icb_long_spans.Rds`. The original coding files are not on the public repository. Public users will load `icb_long_spans.Rds` directly.
    + aligns codings from multiple GUI versions on similar source sentences
 - 02_format_and_clean
    + applies cleaning dictionaries to create `./replication_data/out/ICBe_V1.1_long_clean.Rds`.
 - 03_aggregation
    + applies aggregation algorithm to create `./replication_data/out/ICBe_V1.1_long_agreement.Rds`, `ICBe_V1.1_long.Rds`,  `./replication_data/out/ICBe_V1.1_events_agreed.Rds`, `./replication_data/out/ICBe_V1.1_events_agreed_long.Rds`.
 - 04_validation
    + applies iconography to crises and events to create `ICBe_V1.1_crises_markdown.Rds` and `ICBe_V1.1_events_agreed_markdown.Rds`

### replication_paper

The figures are created in Rmd file for the paper (./replication_paper/pnas_draft/ICBe_pnas_submission_rmd.Rmd). In some cases they have been transformed to other formats via GNU Image Manipulation Program.

  - `case_study_cuban_precision.png`
  - `recall_cuban_and_crimea_andcounts.png`
    + Draws from the [Cuban Missile Automated Case Study googlesheet](https://docs.google.com/spreadsheets/d/1NuRWFB1HEbQbJu_JoEzJ9WYk7kfwwCryF7wBfKPT6uc/edit?usp=sharing) and the [Crimea-Donbas Automated Case Study (redone) googlesheet](https://docs.google.com/spreadsheets/d/1YesAx1CkYCgrEi_WVJ9aho60HesWA6p-3XMma9IzBrI/edit?usp=sharing).
    + uses the iconographry in `./replication_data/in/flags_small/`
  - `p_precision_combined.png`
    + combines metro maps of the two case studies using ICBe_V1.1_events_agreed.Rds
  - `p_semantic_embeddings_dendro.png`
    + plot of semantic embeddings of ICBe_V1.1_events_agreed_markdown.Rds 
  - `p_precision_icews.png`
    + mapping of icews using `./replication_paper/data/out/icews_clean_471_lowest.tsv` (created in `./replication_paper/pnas_draft/appendix.Rmd`)

@@

### External datasets
  - [The ICB project](https://sites.duke.edu/icbdata/) 
    + System-level (icb1v14.csv) and Actor-level (icb2v14.csv) datasets
    + Dyadic-Level Crisis Data ([source](https://sites.duke.edu/icbdata/data-collections/))
  - [Militarized Interstate Disputes (MID)](https://correlatesofwar.org/data-sets/MIDs) version 5.01 at the incident level (MIDI_5.01.Rds) and incident-participant level (MIDIP_5.01.Rds) converted to Rds. 
  - [UCDP Georeferenced Event Dataset (GED) Global version 21.1](https://ucdp.uu.se/downloads/index.html#ged_global) (GEDEvent_v23_1.RData)
  - Cameo Event Codes adapted from [CAMEO Conflict and Mediation Event Observations Codebook](https://parusanalytics.com/eventdata/cameo.dir/CAMEO.09b6.pdf) (cameo.eventcode.txt)
  - [Phoenix Event data](https://databank.illinois.edu/datasets/IDB-2796521)
  - [Terrier event data](https://osf.io/4m2u7/files/) 
    + too large to include in the github repository
    + to replicate, download the folder 'largegeolocatedata' to ICBEdata/replication_paper/data/ignore and decompress
  - [ICEWS data](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/28075&version=30.0)
    + too large to include in the github repository
    + to replicate, download folder 'dataverse_files' to ICBEdata/replication_paper/data/ignore/ and decompress

### Appendix

### 2022-08-10 replication_paper/pnas_draft/ICBe_PNAS_submission_tex.tex
readme update

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

@@

\author{
    Rex W. Douglass
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    Thomas Leo Scherer
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    J. Andrés Gannon
   \\
    Vanderbilt University \\
   \\
  \texttt{} \\
   \And
    Erik Gartzke
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    Jon Lindsay
   \\
    Georgia Institute of Technology \\
   \\
  \texttt{} \\
   \And
    Shannon Carcelli
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Jonathan Wilkenfeld
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    David M. Quinn
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Catherine Aiken
   \\
    Georgetown University \\
   \\
  \texttt{} \\
   \And
    Jose Miguel Cabezas Navarro
   \\
    Universidad Mayor \\
   \\
  \texttt{} \\
   \And
    Neil Lund
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Egle Murauskaite
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Diana Partridge
   \\
    University of Maryland \\
   \\
  \texttt{} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

\usepackage[utf8]{inputenc}
\usepackage{pifont}
\usepackage{newunicodechar}
\newunicodechar{✓}{\ding{51}}
\newunicodechar{✗}{\ding{55}}
\usepackage{array}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{cellspace}
\setlength\cellspacetoplimit{0.8ex}
\renewcommand{\arraystretch}{0.8}
\AtBeginEnvironment{lltable}{\singlespacing}
\usepackage{subcaption}
\usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
\usepackage{float}
\usepackage[algo2e]{algorithm2e}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\begin{document}
\maketitle

\keywords{
    Diplomacy
   \and
    War
   \and
    Crises
   \and
    International Affairs
   \and
    Computational Social Science
  }

\twocolumn

If we could record every important interaction between countries in all
of diplomacy, military conflict, and international political economy,
how much unique information would this chronicle amount to, and how
surprised would we be to see something new? In other words, what is the
entropy of international relations? This record could in principle be
unbounded, but the central conceit of social science is that there are
structural regularities that limit what actors can do, their best
options, and even which actors are likely to survive (1, 2). If so, then
these events can be systematically measured, and accordingly, massive
effort is expended in social science attempting to record these
regularities.\footnote{See work on crises (3, 4), militarized disputes
  (5--7), wars (8, 9), organized violence (10, 11), political violence
  (12), sanctions (13), trade (14), and international agreements
  (15--17), dispute resolution (17, 18), and diplomacy (19, 20).} Thanks
to improvements in natural language processing, more open-ended efforts
have begun to capture entire unstructured streams of events from
international news reports.\footnote{See (21); (22); (23); (24); (25);
  (26). On event-extraction from images and social-media see (27) and
  (28).} How close these efforts are to accurately measuring all or even
most of what is essential in international relations is an open
empirical question, one for which we provide new evidence here.

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{Case Study 1: Cuban Missile Crisis (1962) - ICB Narrative vs. ICBe Events \label{fig:case_study_cuban_precision}}
\centering{\includegraphics[ height=19cm]{"./case_study_cuban_precision.png"}}
\end{figure}
\clearpage
\twocolumn

@@

\clearpage
\onecolumn

{[}1{]} 37 17

@@

The ICBe ontology follows a hierarchical design philosophy where a
smaller number of significant decisions are made early on and then
progressively refined into more specific details (59).\footnote{This
  process quickly focuses the coder on a smaller number of relevant
  options while also allowing them to apply multiple tags if the
  sentence explicitly includes more than one or there is insufficient
  evidence to choose only one tag. The guided coding process also allows
  for the possibility that earlier coarse decisions have less error than
  later fine-grained decisions.} Each coder was instructed to first
thoroughly read the full crisis narrative and then presented with a
custom graphical user interface (SI Appendix, Fig. B1). Coders then
proceeded sentence by sentence, choosing the number of events (0-3) that
occurred, the highest behavior (thought, speech, or activity), a set of
players (\(P\)), whether the means were primarily armed or unarmed,
whether there was an increase or decrease in aggression
(uncooperative/escalating or cooperative/de-escalating), and finally one
or more non-mutually exclusive specific activities. Some additional
details like location and timing information was always collected while
other details were only collected if appropriate, e.g.~force size,
fatalities, domains, units, etc. A unique feature of the ontology is
that thought, speech, and do behaviors can be nested into combinations,
e.g.~an offer for the U.S.S.R. to remove missiles from Cuba in exchange
for the U.S. removing missiles from Turkey. Through compounding, the
ontology can capture what players were said to have known, learned, or
said about other specific fully described actions.

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{Measuring Recall with Synthetic Historical Narratives \label{fig:case_study_cuban_recall}}
\centering{\includegraphics[height=18cm]{./recall_cuban_and_crimea_andcounts.png}}
\textit{Notes: Synthetic narratives combine several thousand accounts of each crisis into a single timeline of events, taking only those mentioned in at least 5 or more documents. Checkmarks represent whether that event could be hand matched to any detail in the ICB corpus, ICBe dataset, or any of the other event datasets.}
\end{figure}
\clearpage
\twocolumn

@@

\clearpage
\onecolumn
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps  \label{fig:p_precision_combined}}
\centering{\includegraphics[width=22cm]{./p_precision_combined.png}}
\end{sidewaysfigure}
\clearpage
\twocolumn

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ICBe event codings in comparison to Semantic Embeddings from source sentences\label{fig:semantic_embeddings}}
\centering{\includegraphics[width=17cm]{./p_semantic_embeddings_dendro.png}}
\textit{Notes: Dots represent individual ICB narrative sentences, as embeded by the Paraphrase-MPNET-base-v2 large language model and flattened into two dimensions with UMAP. Text labels reflect individual leaves of the ICBe ontology, and colors represent intermediate branches of the ontology. Label placement is the median of all of the sentences that tag was applied to by the coders. The dendrogram shows hiearchical clustering of the tags. If ICBe precision is high, the sentences tags were applied to ought to say similar things, and the intended shape of the ontology ought to be visually recognizable. }
\end{figure}
\clearpage
\twocolumn

@@

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ ICEWs Events by Day by Type during the Crimea-Donbas Crisis \label{fig:p_precision_icews}}
\centering{\includegraphics[width=17cm]{./p_precision_icews.png}}
\textit{Notes:Unit of analysis is the Dyad-Day. Edges <-> indicates undirected dyad and -> indicates directed dyad. Top 10 most active dyads per category shown. Red text shows events from the synthetic narrative relative to that event category. Blue bars indicate an event recorded by ICEWs for that dyad on that day. }
\end{figure}
\clearpage
\twocolumn

@@

\hypertarget{format}{%
\section*{Acknowledgements}\label{format}}
\addcontentsline{toc}{section}{Acknowledgements}

We thank the ICB Project and its directors and contributors for their
foundational work and their help with this effort. We make special
acknowledgment of Michael Brecher for helping found the ICB project in
1975, creating a resource that continues to spark new insights to this
day. We thank the many undergraduate coders for their patience and
dedication. Thanks to the Center for Peace and Security Studies and its
membership for comments. Special thanks to Rebecca Cordell, Philip
Schrodt, Zachary Steinert-Threlkeld, and Zhanna Terechshenko for
generous feedback. Thank you to the cPASS research assistants that
contributed to this project: Helen Chung, Daman Heer, Syeda ShahBano
Ijaz, Anthony Limon, Erin Ling, Ari Michelson, Prithviraj Pahwa, Gianna
Pedro, Tobias Stodiek, Yiyi `Effie' Sun, Erin Werner, Lisa Yen, and
Ruixuan Zhang. This project was supported by a grant from the Office of
Naval Research {[}N00014-19-1-2491{]} and benefited from the Charles
Koch Foundation's support for the Center for Peace and Security Studies.

\hypertarget{format}{%
\section*{Author Contributions}\label{format}}
\addcontentsline{toc}{section}{Author Contributions}

Conceptualization: R.W.D., E.G., J.L.; Methodology: R.W.D., T.L.S.;
Software: R.W.D.; Validation: R.W.D., T.L.S.; Formal Analysis: R.W.D.,
T.L.S.; Investigation: S.C., R.W.D., J.A.G., C.K., N.L., E.M., J.M.C.N.,
D.P., D.Q., J.W.; Data Curation: R.W.D., D.Q., T.L.S., J.W.; Writing -
Original Draft: R.W.D., T.L.S.; Writing - Review \& Editing: R.W.D.,
J.A.G., E.G., T.L.S.; Visualization: R.W.D., T.L.S.; Supervision: E.G.;
Project Administration: S.C., R.W.D., J.A.G., D.Q., T.L.S., J.W.;
Funding Acquisition: E.G., J.L.

@@

\bibliographystyle{unsrt}
\bibliography{paper.bib}

>>>>

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

@@

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Rex W. Douglass}
\author[a]{Thomas Leo Scherer} 
\author[b]{J. Andrés Gannon}
\author[a]{Erik Gartzke}
\author[c]{Jon Lindsay}
\author[d]{Shannon Carcelli}
\author[d]{Jonathan Wilkenfeld}
\author[d]{David M. Quinn}
\author[e]{Catherine Aiken}
\author[f]{Jose Miguel Cabezas Navarro}
\author[d]{Neil Lund}
\author[d]{Egle Murauskaite}
\author[d]{Diana Partridge}

\affil[a]{University of California, San Diego }
\affil[b]{Vanderbilt University}
\affil[c]{Georgia Institute of Technology}
\affil[d]{University of Maryland}
\affil[e]{Georgetown University}
\affil[f]{Universidad Mayor}

% Please give the surname of the lead author for the running footer
\leadauthor{Douglass}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Countries routinely face crises that risk escalating into full scale war but we do not have systematic measurements of the progression of past crises and what moves and counter moves led to or helped avoid war. Instead policy makers typically rely on one or two historical analogies,
chosen through ad hoc selection criteria, and described in unsystematic terms. This paper introduces a new scientific approach to measuring the step by step moves of international crises over the last hundred years, combining subject expertise with state of the art natural language processing and machine learning methods. It serves as a guide for constructing and evaluating large scale measurement collection in the social sciences.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Conceptualization: R.W.D., E.G., J.L.; Methodology: R.W.D., T.L.S.; Software: R.W.D.; Validation: R.W.D., T.L.S.; Formal Analysis: R.W.D., T.L.S.; Investigation: S.C., R.W.D., J.A.G., C.K., N.L., E.M., J.M.C.N., D.P., D.Q., J.W.; Data Curation: R.W.D., D.Q., T.L.S., J.W.; Writing -
Original Draft: R.W.D., T.L.S.; Writing - Review \& Editing: R.W.D., J.A.G., E.G., T.L.S.; Visualization: R.W.D., T.L.S.; Supervision: E.G.; Project Administration: S.C., R.W.D., J.A.G., D.Q., T.L.S., J.W.; Funding Acquisition: E.G., J.L.}
% \authordeclaration{Please declare any conflict of interest here.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: rexdouglass\@gmail.com }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Diplomacy $|$ War $|$ Crises $|$ International Affairs $|$ Computational Social Science}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX TO BE ADDED}}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

\usepackage[utf8]{inputenc}
\usepackage{pifont}
\usepackage{newunicodechar}
\newunicodechar{✓}{\ding{51}}
\newunicodechar{✗}{\ding{55}}
\usepackage{array}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{cellspace}
\setlength\cellspacetoplimit{0.8ex}
\renewcommand{\arraystretch}{0.8}
\AtBeginEnvironment{lltable}{\singlespacing}
\usepackage{subcaption}
\usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
\usepackage{float}
\usepackage[algo2e]{algorithm2e}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\dropcap{I}f we could record every important interaction between countries in all
of diplomacy, military conflict, and international political economy,
how much unique information would this chronicle amount to, and how
surprised would we be to see something new? In other words, what is the
entropy of international relations? This record could in principle be
unbounded, but the central conceit of social science is that there are
structural regularities that limit what actors can do, their best
options, and even which actors are likely to survive (1, 2). If so, then
these events can be systematically measured, and accordingly, massive
effort is expended in social science attempting to record these
regularities.\footnote{See work on crises (3, 4), militarized disputes
  (5--7), wars (8, 9), organized violence (10, 11), political violence
  (12), sanctions (13), trade (14), and international agreements
  (15--17), dispute resolution (17, 18), and diplomacy (19, 20).} Thanks
to improvements in natural language processing, more open-ended efforts
have begun to capture entire unstructured streams of events from
international news reports.\footnote{See (21); (22); (23); (24); (25);
  (26). On event-extraction from images and social-media see (27) and
  (28).} How close these efforts are to accurately measuring all or even
most of what is essential in international relations is an open
empirical question, one for which we provide new evidence here.

@@

\begin{figure*}
\caption{Case Study 1: Cuban Missile Crisis (1962) - ICB Narrative vs. ICBe Events \label{fig:case_study_cuban_precision}}
\centering{\includegraphics[height=19cm]{"./case_study_cuban_precision.png"}}
\end{figure*}

@@

The ICBe ontology follows a hierarchical design philosophy where a
smaller number of significant decisions are made early on and then
progressively refined into more specific details (59).\footnote{This
  process quickly focuses the coder on a smaller number of relevant
  options while also allowing them to apply multiple tags if the
  sentence explicitly includes more than one or there is insufficient
  evidence to choose only one tag. The guided coding process also allows
  for the possibility that earlier coarse decisions have less error than
  later fine-grained decisions.} Each coder was instructed to first
thoroughly read the full crisis narrative and then presented with a
custom graphical user interface (SI Appendix, Fig. B1). Coders then
proceeded sentence by sentence, choosing the number of events (0-3) that
occurred, the highest behavior (thought, speech, or activity), a set of
players (\(P\)), whether the means were primarily armed or unarmed,
whether there was an increase or decrease in aggression
(uncooperative/escalating or cooperative/de-escalating), and finally one
or more non-mutually exclusive specific activities. Some additional
details like location and timing information was always collected while
other details were only collected if appropriate, e.g.~force size,
fatalities, domains, units, etc. A unique feature of the ontology is
that thought, speech, and do behaviors can be nested into combinations,
e.g.~an offer for the U.S.S.R. to remove missiles from Cuba in exchange for the U.S. removing \linebreak \clearpage
\onecolumn

% {[}1{]} 37 17

@@

missiles from Turkey. Through compounding, the ontology 
can capture what players were said to have known, learned, or
said about other specific fully described actions.

@@

\begin{figure*}
\caption{Measuring Recall with Synthetic Historical Narratives \label{fig:case_study_cuban_recall}}
\centering{\includegraphics[height=18cm]{./recall_cuban_and_crimea_andcounts.png}}
\textit{Notes: Synthetic narratives combine several thousand accounts of each crisis into a single timeline of events, taking only those mentioned in at least 5 or more documents. Checkmarks represent whether that event could be hand matched to any detail in the ICB corpus, ICBe dataset, or any of the other event datasets.}
\end{figure*}

@@

\begin{sidewaysfigure*}
\caption{Crisis Maps  \label{fig:p_precision_combined}}
\centering{\includegraphics[width=22cm]{./p_precision_combined.png}}
\end{sidewaysfigure*}

@@

\begin{figure*}
\caption{ICBe event codings in comparison to Semantic Embeddings from source sentences\label{fig:semantic_embeddings}}
\centering{\includegraphics[width=17cm]{./p_semantic_embeddings_dendro.png}}
\textit{Notes: Dots represent individual ICB narrative sentences, as embeded by the Paraphrase-MPNET-base-v2 large language model and flattened into two dimensions with UMAP. Text labels reflect individual leaves of the ICBe ontology, and colors represent intermediate branches of the ontology. Label placement is the median of all of the sentences that tag was applied to by the coders. The dendrogram shows hiearchical clustering of the tags. If ICBe precision is high, the sentences tags were applied to ought to say similar things, and the intended shape of the ontology ought to be visually recognizable. }
\end{figure*}

@@

\begin{figure*}
\caption{ ICEWs Events by Day by Type during the Crimea-Donbas Crisis \label{fig:p_precision_icews}}
\centering{\includegraphics[width=17cm]{./p_precision_icews.png}}
\textit{Notes:Unit of analysis is the Dyad-Day. Edges <-> indicates undirected dyad and -> indicates directed dyad. Top 10 most active dyads per category shown. Red text shows events from the synthetic narrative relative to that event category. Blue bars indicate an event recorded by ICEWs for that dyad on that day. }
\end{figure*}

@@

@@

\matmethods{This paper's data, online appendix, and replication material as well as visualizations of every historical episode available at a companion website www.crisisevents.org and the github repository \href{https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$}{ICBEventData}.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{We thank the ICB Project and its directors and contributors for their
foundational work and their help with this effort. We make special
acknowledgment of Michael Brecher for helping found the ICB project in
1975, creating a resource that continues to spark new insights to this
day. We thank the many undergraduate coders for their patience and
dedication. Thanks to the Center for Peace and Security Studies and its
membership for comments. Special thanks to Rebecca Cordell, Philip
Schrodt, Zachary Steinert-Threlkeld, and Zhanna Terechshenko for
generous feedback. Thank you to the cPASS research assistants that
contributed to this project: Helen Chung, Daman Heer, Syeda ShahBano
Ijaz, Anthony Limon, Erin Ling, Ari Michelson, Prithviraj Pahwa, Gianna
Pedro, Tobias Stodiek, Yiyi `Effie' Sun, Erin Werner, Lisa Yen, and
Ruixuan Zhang. This project was supported by a grant from the Office of
Naval Research {[}N00014-19-1-2491{]} and benefited from the Charles
Koch Foundation's support for the Center for Peace and Security Studies.}

\showacknow{} % Display the acknowledgments section

% Bibliography

@@

\bibliographystyle{unsrt}

\bibliography{paper.bib}
